---
title: "Ch 8 - Soap Production Lines Example"
author: "Adam McQuistan"
date: "Saturday, April 30, 2016"
output: pdf_document
---


# The dataset

| Variable | Description                                             |
|----------|---------------------------------------------------------|
| Y        | Amount of scrap produced                                |
| X1       | Quantitative variable. Line Speed                       |
| X2       | Qualitative variable. 1 for line 1, 0 for line 2        |

```{r, warning=FALSE, message=FALSE}
lines <- readLines("CH08TA05.txt")
Y <- vector()
X1 <- vector()
X2 <- vector()

library(gdata)
for(line in lines){
  line <- trim(line)
  lineAry <- unlist(strsplit(line, "  "))
  Y <- c(Y, lineAry[1])
  X1 <- c(X1, lineAry[2])
  X2 <- c(X2, lineAry[3])
}

df <- data.frame(Y=as.numeric(Y),
                 X1=as.numeric(X1), 
                 X2=factor(X2, levels=c(0,1), labels=c("Line2", "Line1")))

str(df)
summary(df)


library(car)
scatterplot(Y ~ X1 | X2, data=df,
            ylab="Scrap",
            xlab="Line Speed")
```



__Tenative Model__

$$Y = \beta_0 + \beta_1 X_1 + \beta_2 X_2 + \beta_3 X_1 X_2 + \varepsilon$$

```{r, echo=FALSE}
fullRegressionAnova <- function(lm_anova){
  VariationSource <- c("Regression", rownames(lm_anova), "Total")
  SSR <- sum(lm_anova$"Sum Sq"[1:(length(lm_anova$"Sum Sq")-1)])
  SST <- sum(lm_anova$"Sum Sq")
  DFReg <- sum(lm_anova$"Df"[1:(length(lm_anova$"Df")-1)])
  MSE <- lm_anova$"Mean Sq"[length(lm_anova$"Mean Sq")]
  MSR <- SSR / DFReg  
  
  SS <- c(SSR,lm_anova$"Sum Sq",SST)
  MS <- c(MSR, lm_anova$"Mean Sq", NA)
  DF <- c(DFReg, lm_anova$"Df", sum(lm_anova$"Df"))
  F_stat <- MSR / MSE
  F_stats <-c(F_stat, lm_anova$"F value",NA)
  df_out <- data.frame(VariationSource, DF,SS, MS, F_stats)
  print(df_out)
  return(df_out)
}
```

```{r}
result1 <- lm(Y ~ X1 + X2 + I(X1 * as.numeric(X2)), data=df)
result1_smry <- summary(result1)
df$residuals <- result1_smry$residuals
print(result1_smry)

result1_aov <- fullRegressionAnova(anova(result1))
```

# Brown-Forsythe Test for Equal Error Variance in the Lines

$H_o$: equal variances

$H_a$: the two variances are not equal

```{r}
leveneTest(df$residuals, df$X2, center="median")
```

Since pvalue > 0.05 conclude equal variances which leads us to conclude the data is approriate for the regression model. 

# Inferences about Two Regression Lines

$H_o: \beta_2 = \beta_3 = 0$

$H_a:$ Not both $\beta_2 = 0$ and $\beta_3 = 0$

```{r}
term1 <- (result1_aov$SS[3] + result1_aov$SS[4]) / (result1_aov$DF[3] + result1_aov$DF[4])
term2 <- result1_aov$SS[5] / result1_aov$DF[5]
F_stat <- term1 / term2

F_stat
F_crit <- qf(0.99, 2,23)

msg <- paste("F stat: ", F_stat, "\nF crit: ", F_crit, sep="")
result <- ifelse(F_stat > F_crit, "Conclude Ha, not identical", "Conclude Ho, they are identical")
cat(msg, "\n", result, sep="")

```