---
title: "Ch 6 Example - Dwaine Studio Sales"
author: "Adam McQuistan"
date: "Sunday, April 03, 2016"
output: pdf_document
---


```{r, warning=FALSE}
library("ggplot2")
setwd("C:\\Users\\AdamMcQuistan\\Documents\\ISQA 8340\\ch 6")
df <- read.csv("data/ch 6 sales.csv")
df$id = NULL
df <- with(df, data.frame(SALES, DispIncome, TargetPop))

```


##View General Linearity Plots Among Predictors and Outcome Vars

```{r}
pairs(df, pch=21)
```

Sales appears to have a linear relationship with the target population (individuals less than or equal to 16 yrs old) and the disploable income of the cities.  


##Building the Model

```{r}
result <- lm(SALES ~ DispIncome + TargetPop, data=df)
result_smry <- summary(result)
F_stat <- round(as.numeric(result_smry$fstatistic["value"]),1)
F_crit <- round(qf(0.95, df1=2, df2=result_smry$df[2]),1)
result_smry
```

Analyze and test for the relatedness of sales to the target population and per capita disposable income.
F critical is `r  F_crit`.

$H_0: \beta_1 = 0$ and $\beta_2 = 0$ for F* <= F critical

$H_a:$ not both $\beta_1$ and $\beta_2$ equal zero for F* > F critical

Since F* ( `r F_stat` ) > F critical ( `r F_crit` ) we conclude that sales are related to target population and per capital income but, the usefullness of the model for making predictions have yet to be determined.


The P-value is < 0.05 further indicating the rejection of $H_o$. 



##Part C. Calculation coefficient of multople determination.

The Multiple $R^2$ value is 0.9167 which translates into the model explaining about 92% of the variation in sales.  The Adjusted $R^2$ value of 0.9075 shows the correction for more than on predictor variable overfitting which is a very small change from the standard multiple r-squared value indicating little effects of overfitting. 

To analyze and test the individual parameters of the regression model we use the students t-test such that:

$H_0: \beta_p = 0$ indicating no relationship 

$H_a:$ $\beta_p \neq 0$ indicating there is a relationship

Since the p-values for both parameters (Disposable income and Target population) are less than 0.05 we conclude each individual parameter is significant and related to sales. 

##Graphical Assessment of the Model (Residual Analysis)

```{r}
df_model <- result$model[,1:3]
df_model$FittedVals <- result$fitted.values
df_model$Residuals <- result$residuals

plot(x=df_model$FittedVals, y=df_model$Residuals, xlab="Fitted Values", ylab="Residuals")

plot(x=df_model$TargetPop, y=df_model$Residuals, xlab="Target Population", ylab="Residuals")

plot(x=df_model$DispIncome, y=df_model$Residuals, xlab="Disposable Income", ylab="Residuals")
```
