---
title: "Exam 3 - Question 1"
author: "Adam McQuistan"
date: "Saturday, April 30, 2016"
output: pdf_document
---

# Problem 1 - 8.24

The dataset

| Variable | Description                                             |
|----------|---------------------------------------------------------|
| Price    | Selling price of house                                  |
| Assessed | Quantitative variable. Assessed value in thousands USD  |
| Lot      | Qualitative variable. 1 for corner lot, 0 for non corner|

```{r}
df <- read.csv(file="data/8.24.csv")
df$Case <- NULL
df$Lot <- factor(df$Lot, levels=c(0,1), labels=c("Non-Corner", "Corner"))
summary(df)
```

## Part A

Plot the sample data for the two populations as a symbolic scatter plot. Does the regression appear to be the same for the two populations?

```{r, warning=FALSE, message=FALSE}
library(car)
scatterplot(Price ~ Assessed | Lot, data=df,
            ylab="Selling Price ($1000s)",
            xlab="Assessed Value ($1000s)")
```


Yes they are appear to be different.  The slope for non-corner lots is greater and, on average the non-corner lots tend to sell above the corner lots for the same assessed values.  


_Note that the zero symbol is for non-corner lots and triangles are for corner lots._


# Part B

Test for identity of the regression functuons for dwellings on corner and non-corner lots. Control for risk of type 1 error at 0.05. 

$$Y = \beta_0 + \beta_1 X_1 + \beta_2 X_2 + \beta_3 X_1 X_2 + \varepsilon$$

```{r, echo=FALSE}
fullRegressionAnova <- function(lm_anova){
  VariationSource <- c("Regression", rownames(lm_anova), "Total")
  SSR <- sum(lm_anova$"Sum Sq"[1:(length(lm_anova$"Sum Sq")-1)])
  SST <- sum(lm_anova$"Sum Sq")
  DFReg <- sum(lm_anova$"Df"[1:(length(lm_anova$"Df")-1)])
  MSE <- lm_anova$"Mean Sq"[length(lm_anova$"Mean Sq")]
  MSR <- SSR / DFReg  
  
  SS <- c(SSR,lm_anova$"Sum Sq",SST)
  MS <- c(MSR, lm_anova$"Mean Sq", NA)
  DF <- c(DFReg, lm_anova$"Df", sum(lm_anova$"Df"))
  F_stat <- MSR / MSE
  F_stats <-c(F_stat, lm_anova$"F value",NA)
  df_out <- data.frame(VariationSource, DF,SS, MS, F_stats)
  print(df_out)
  return(df_out)
}
```

```{r}
result1 <- lm(Price ~ Assessed + Lot + I(Assessed * as.numeric(Lot)),
              data=df)
result1_smry <- summary(result1)
df$residuals <- result1_smry$residuals
print(result1_smry)

result1_aov <- fullRegressionAnova(anova(result1))
```


### Inferences about Two Regression Lines

$H_o: \beta_2 = \beta_3 = 0$

$H_a:$ Not both $\beta_2 = 0$ and $\beta_3 = 0$

```{r}
term1 <- (result1_aov$SS[3] + result1_aov$SS[4]) / (result1_aov$DF[3] + result1_aov$DF[4])
term2 <- result1_aov$SS[5] / result1_aov$DF[5]
F_stat <- term1 / term2
F_crit <- qf(0.95, 2,60)

msg <- paste("F stat: ", F_stat, "\nF crit: ", F_crit, sep="")
result <- ifelse(F_stat > F_crit, "Conclude Ha, not identical", "Conclude Ho, they are identical")
cat(msg, "\n", result, sep="")
```

Since F stat (18.7) > F crit (F(0.95, 2, 60) = 3.2) we conclude Ha that the regression functions are not identical.

# Part C

Plot the estimated regression function for the two populations and describe the nature of the differences between them. 

```{r}
scatterplot(Price ~ Assessed | Lot, data=df,
            ylab="Selling Price ($1000s)",
            xlab="Assessed Value ($1000s)")
```

The slope for non-corner lots is greater and, on average the non-corner lots tend to sell above the corner lots for the same assessed values.