---
title: "Ch 9 - Surgical Unit Example"
author: "Adam McQuistan"
date: "Sunday, May 01, 2016"
output: pdf_document
---

# The dataset

| Variable | Description                                             |
|----------|---------------------------------------------------------| 
| X1       | blood clotting score                                    |
| X2       | prognostic index                                        |
| X3       | enzyme function score                                   |
| X4       | liver function test score                               |
| X5       | age in years                                            |
| X6       | gender, 0 for male and 1 for female                     |
| X7, X8   | alcohol use. see table below                            |
| Y        | Survival time                                           |


| Alcohol Use | X7 | X8 |
|-------------|----|----|
| None        | 0  | 0  |
| Moderate    | 1  | 0  |
| Severe      | 0  | 1  |


```{r}
df <- read.table(file="CH09TA01.txt", sep="\t", header=F)
names(df) = c("X1","X2","X3","X4","X5","X6","X7","X8","Y", "lnY")
str(df)
```

## Subsetting the Data into First 54 Cases and First 4 Variables

```{r}
df54 <- df[1:54,c("X1","X2","X3","X4","Y","lnY")]
resultY <- lm(Y ~ X1 + X2 + X3 + X4, data=df)
resultlnY <- lm(lnY ~ X1 + X2 + X3 + X4, data=df)
```

## Residual Plots of Y and lnY

```{r, fig.height=3.5, fig.width=6, echo=FALSE}
df54$ResidualsY <- resultY$residuals; df54$ResidualslnY <- resultlnY$residuals
df54$PredictedValsY <- resultY$fitted.values; df54$PredictedValslnY <- resultlnY$fitted.values

with(df54, {
     plot(x=PredictedValsY, y=ResidualsY,
          ylim=c(-max(ResidualsY), max(ResidualsY)),
          xlab="Predicted Values for Y", ylab="Residuals for Y", main="")
     points(c(min(PredictedValsY), max(PredictedValsY)), 
            c(0,0), type="l", lwd="2", col="blue")})

with(df54, {
     plot(x=PredictedValslnY, y=ResidualslnY,
          ylim=c(-max(ResidualslnY), max(ResidualslnY)),
          xlab="Predicted Values for ln(Y)", ylab="Residuals for ln(Y)", main="")
     points(c(min(PredictedValslnY), max(PredictedValslnY)), 
            c(0,0), type="l", lwd="2", col="blue")})
```

```{r, fig.height=3.6, fig.width=6, echo=FALSE}
qqnorm(resultY$residuals, ylab="Residuals Y", main=""); qqline(resultY$residuals)

qqnorm(resultlnY$residuals, ylab="Residuals ln(Y)", main=""); qqline(resultlnY$residuals)
```

# Correlations and Scatter Plot Matrix

```{r}
with(df54, cor(df54[,c("lnY","X1","X2","X3","X4")]))
```

```{r}
with(df54, pairs(df54[,c("lnY","X1","X2","X3","X4")]))
```

# Using $R_{p}^{2}$ and $SSE_p$

```{r, warning=FALSE, message=FALSE}
library(leaps)

evaluateRegressionModel <- function(x, y, method, names){
  result <- leaps(x=x, y=y,method=method,names=names)
  labels <- result$label[2:length(result$label)]
  
  Variables <- vector()
  VariablesCnt <- vector()
  metric <- vector()
  
  for(rowIdx in 1:dim(result$which)[1]){
    selected <- result$which[rowIdx,]
    VariablesCnt <- c(VariablesCnt, sum(result$which[rowIdx,]))
    vars <- paste(labels[selected], collapse=" ")
    Variables <- c(Variables, vars)  
    
    thisMetric <- switch(method,
                     r2=result$r2[rowIdx],
                     Cp=result$Cp[rowIdx],
                     adjr2=result$adjr2[rowIdx])
              
    metric <- c(metric, thisMetric)
  }
  
  out <- data.frame(Variables, VariablesCnt, metric)
  names(out)[3] = method
  print(out)
  return(out)
}

result <- evaluateRegressionModel(x=as.matrix(df54[,1:4]),
                                  y=df54$lnY,
                                  method="r2",
                                  names=names(df54)[1:4])

library(dplyr)
df_tbl <- tbl_df(result) %>% 
            group_by(VariablesCnt) %>% 
            summarize(MaxR2 = max(r2))

plot(x=result$VariablesCnt, y=result$r2)
lines(df_tbl$VariablesCnt, df_tbl$MaxR2)
```

# Mallows $C_p$ Criterion

```{r}
result <- evaluateRegressionModel(x=as.matrix(df54[,1:4]),
      y=df54$lnY,
      method="Cp",
      names=names(df54)[1:4])

df_tbl <- tbl_df(result) %>% 
            group_by(VariablesCnt) %>% 
            summarize(MinCp = min(Cp))

plot(x=result$VariablesCnt, y=result$Cp)
lines(df_tbl$VariablesCnt, df_tbl$MinCp)
```


# Stepwise Regression - Forward

For forward stepwise regression it is important to identify an $\alpha$ cut off for determining which predictors to let into the model. For example, if your cut of is 0.05 then you would only include variables with pvalues below the variable. 

```{r, warning=FALSE, message=FALSE}
library(MASS)

full <- lm(lnY ~ X1 + X2 + X3 + X4 + X5 + X6 + X7 + X8, 
           data=df)
Null <- lm(lnY ~ 1, data=df)
addterm(Null, scope=full, test="F")
```

For this iteration we would include the predictor with the lowest p value which is X3.
A new base model is built by including the X3 value then the procedure is ran again. 

```{r}
newModel <- lm(lnY ~ X3, data=df)
addterm(newModel, scope=full, test="F")
```

Now I'll add in X2


```{r}
newModel <- lm(lnY ~ X3 + X2, data=df)
addterm(newModel, scope=full, test="F")
```

This time X1 is added


```{r}
newModel <- lm(lnY ~ X3 + X2 + X1, data=df)
addterm(newModel, scope=full, test="F")
```

This time through the only variable making the cut off is X8. Adding this gives us our full model. 


```{r}
newModel <- lm(lnY ~ X3 + X2 + X1 + X8, data=df)
addterm(newModel, scope=full, test="F")
```

# PRESS Criterion

```{r}
PRESS <- function(linear.model) {
  pr <- residuals(linear.model) / (1-lm.influence(linear.model)$hat)
  press <- sum(pr^2)
  print(press)
  return(press)
}
```

# Model Validation 

Three Models are Evaluated

$lnY = \beta_0 + \beta_1 X_1 + \beta_2 X_2 + \beta_3 X_3 + \beta_8 X_8 + \varepsilon$

$lnY = \beta_0 + \beta_1 X_1 + \beta_2 X_2 + \beta_3 X_3 + \beta_5 X_5 + \beta_8 X_8 + \varepsilon$

$lnY = \beta_0 + \beta_1 X_1 + \beta_2 X_2 + \beta_3 X_3 + \beta_5 X_5 + \beta_6 X_6 + \beta_8 X_8 + \varepsilon$


```{r, echo=FALSE}
fullRegressionAnova <- function(lm_anova){
  VariationSource <- c("Regression", rownames(lm_anova), "Total")
  SSR <- sum(lm_anova$"Sum Sq"[1:(length(lm_anova$"Sum Sq")-1)])
  SST <- sum(lm_anova$"Sum Sq")
  DFReg <- sum(lm_anova$"Df"[1:(length(lm_anova$"Df")-1)])
  MSE <- lm_anova$"Mean Sq"[length(lm_anova$"Mean Sq")]
  MSR <- SSR / DFReg  
  
  SS <- c(SSR,lm_anova$"Sum Sq",SST)
  MS <- c(MSR, lm_anova$"Mean Sq", NA)
  DF <- c(DFReg, lm_anova$"Df", sum(lm_anova$"Df"))
  F_stat <- MSR / MSE
  F_stats <-c(F_stat, lm_anova$"F value",NA)
  df_out <- data.frame(VariationSource, DF,SS, MS, F_stats)
  print(df_out)
  return(df_out)
}


```


```{r}
dfTrain <- df[,c("lnY", "X1","X2","X3","X5","X6","X8")]
result1 <- lm(lnY ~ X1 + X2 + X3 + X8, data=df)
result2 <- lm(lnY ~ X1 + X2 + X3 + X5 + X8, data=df)
result3 <- lm(lnY ~ X1 + X2 + X3 + X5 + X6 + X8, data=df)

result1_aov <- fullRegressionAnova(anova(result1))
result2_aov <- fullRegressionAnova(anova(result2))
result3_aov <- fullRegressionAnova(anova(result3))

statistic <- c("p","SSEp","PRESSp","Cp","MSEp","R2p")
model1 <- rep(NA, times=length(statistic))
model2 <- rep(NA, times=length(statistic))
model3 <- rep(NA, times=length(statistic))

model1[1] <- result1_aov$DF[1] + 1
model2[1] <- result2_aov$DF[1] + 1
model3[1] <- result3_aov$DF[1] + 1

model1[2] <- result1_aov$SS[(length(result1_aov$SS)-1)]
model2[2] <- result2_aov$SS[(length(result2_aov$SS)-1)]
model3[2] <- result3_aov$SS[(length(result3_aov$SS)-1)] 

model1[3] <- PRESS(result1)
model2[3] <- PRESS(result2)
model3[3] <- PRESS(result3)



model1[5] <- result1_aov$MS[(length(result1_aov$MS)-1)]
model2[5] <- result2_aov$MS[(length(result2_aov$MS)-1)]
model3[5] <- result3_aov$MS[(length(result3_aov$MS)-1)] 



```
